---
title: "NMR Findings, Additional"
author: "Zach A. Bright"
date: "5/4/2021"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plotly)
library(phyloseq)
library(devtools)
library(ggbiplot)
library("dplyr")
library(vegan)
library(ggplot2)

# Import and Format the Data
# Load NMR data. 
nmr.data <- read.csv("nmr_data.csv")
# nmr.data$Sample.ID <- as.numeric(sub("A", "", nmr.data$Sample.ID))
# Load meta data
meta.data <- read.csv("sample_metadata.csv")
# Extract only the NMR data samples.
meta.data <- subset(meta.data, Sample.Description == "Primary" & PMAxx. == "")
# meta.data <- meta.data[meta.data$Sample.Description == "Primary" && meta.data$PMAxx., ]
# Merge the dataframes.
nmr.meta <- merge(meta.data, nmr.data)
# Sample IDs must match those of other files. 
nmr.meta$Sample.ID <- sub("A", "", nmr.meta$Sample.ID)
nmr.meta$Sample.ID <- sub("^", "HA", nmr.meta$Sample.ID)
rownames(nmr.meta) <- nmr.meta$Sample.ID
nmr.meta["Age..days."] <- as.numeric(nmr.meta$Age..days.)
```

## NMR-detected molecules
```{r phyloseq setup, include=FALSE}
# These phyloseq objects were made by Keisha Rose Harrison.
dfbacteria <- readRDS("/home/zachary/Documents/Classes/MB 599/phyloseq objects/dfbacteria.rds")
dffungi <- readRDS("/home/zachary/Documents/Classes/MB 599/phyloseq objects/dffungi.rds")
```

Create a function to automatically adjust p-values for linear models. Neither of these is actually used. 
```{r, include = FALSE}
# Note: We need a function that does this for multiple linear models. 
adjust_p_values <- function(model_summary) {
  pvals <- model_summary$coefficients[, 4]
  model_summary$coefficients[, 4] <- p.adjust(pvals)
  return(model_summary)
}
# This function is actually junk. 
adjust_p_values_multiple <- function(model_list) {
  pvals <- c()
  for(model in model_list) {
    pvals <- append(pvals, model$coefficients[, 4])
  }
  pvals_adj <- p.adjust(pvals)
  for(model in model_list) {
    n <- length(model$coefficients[, 4])
    model$coefficient[, 4] <- pvals_adj[1:n]
    pvals_adj < pvals_adj[n:length(pvals)]
  }
  return(model_list)
}
```

### Fungi
Same as in NMR_Findings_Initial, but ignoring more uninteresting chemicals. 
```{r}
# OTU table for fungi
# Only worry about genera
dffungi <- tax_glom(dffungi, taxrank = "Genus")

otu.fungi <- otu_table(dffungi)
otu.fungi <- otu.fungi[, intersect(rownames(nmr.meta), colnames(otu.fungi))]
dim(otu.fungi)

otu.fungi <- sweep(otu.fungi, 2, colSums(otu.fungi), "/")
# Require that at each taxon is present at at least 5% in at least 3 samples
otu.fungi <- t(na.exclude(otu.fungi[rowSums(otu.fungi > 0.001) > 2, ]))
nmr.meta.fungi <- nmr.meta[intersect(rownames(nmr.meta), rownames(otu.fungi)), 15:42]
nmr.meta.fungi <- subset(nmr.meta.fungi, select = c(
  -X2.Hydroxyisobutyrate,
  -X2.Hydroxyisovalerate,
  -X4.Guanidinobutanoate,
  -Caffeine,
  -Glucuronic.acid,
  -Glycerol,
  -Citraconate
    )
  )
mean_normalize <- function(vec) {
  return(vec / mean(vec))
}
nmr.meta.fungi[, 2:dim(nmr.meta.fungi)[2]] <- apply(
  nmr.meta.fungi[, 2:dim(nmr.meta.fungi)[2]], 
  2, 
  mean_normalize
  )
```

For each taxon, make a model using all the chemicals as explanatory variables, then use stepwise regression to find a stronger model. This code automatically does FDR correction for the variables included, but DOES NOT apply necessary corrections for the model selection process. 
```{r}
tax_df <- data.frame(tax_table(dffungi))

pvals <- c()
for (species in colnames(otu.fungi)) {
  model.initial <- lm(otu.fungi[, species] ~ Age..days. : ., data = nmr.meta.fungi)
  #model.final <- summary(model.initial)
  model.final <- summary(step(model.initial, trace = FALSE))
  pvals <- append(pvals, model.final$coefficients[, 4])
  #adjust_p_values_multiple(models)
}
pvals_adj <- p.adjust(pvals)
for (species in colnames(otu.fungi)) {
  model.initial <- lm(otu.fungi[, species] ~ Age..days. : ., data = nmr.meta.fungi)
  model.final <- summary(step(model.initial, trace = FALSE))
  #model.final <- summary(model.initial)
  n <- length(model.final$coefficients[, 4])
  model.final$coefficients[, 4] <- pvals_adj[1:n]
  pvals_adj < pvals_adj[n + 1:length(pvals)]
  print(paste(species, " (", tax_df[species, "Genus"], ")", ":", sep = ""))
  print(model.final)
  #print(pvals_adj)
  #print(n)
}

```

Create a linear model using one chemical as the response variable and all others as explanatory variables. Then plot the residuals vs. some species of interest. 
```{r}
plot_resid_fungi <- function(taxon, chemical){
  tax_df <- data.frame(tax_table(dffungi))
  genus <- tax_df[taxon, "Genus"]
  lm(nmr.meta.fungi[, chemical] ~ ., data = nmr.meta.fungi[colnames(nmr.meta.fungi) != chemical]) -> lm1
  step(lm1, trace = FALSE) -> lm1
  plot_title <- paste(genus, "vs", chemical, "residuals")

    ggplot(mapping = aes(
    otu.fungi[, taxon],
    lm1$residuals)) +
    geom_point() +
    geom_smooth(method = "lm") +
    xlab(taxon) +
    ylab(chemical) +
    ggtitle(plot_title)

    
    if (askYesNo("Save plot?")){
      ggsave(paste("~/Documents/Classes/MB 599/Plots/", plot_title, ".png", sep = ""))
  }
}

```

Create a simple plot of a fungal species vs a chemical. 
```{r}
plot_fungi <- function(taxon, chemical){
  tax_df <- data.frame(tax_table(dffungi))
  genus <- tax_df[taxon, "Genus"]
  plot_title <- paste(genus, "vs", chemical)

    ggplot(mapping = aes(
    otu.fungi[, taxon],
    nmr.meta.fungi[, chemical])) +
    geom_point() +
    geom_smooth(method = "lm") +
    xlab(taxon) +
    ylab(chemical) +
    ggtitle(plot_title)
  
  if (askYesNo("Save plot?")){
    ggsave(paste("~/Documents/Classes/MB 599/Plots/", plot_title, ".png", sep = ""))
  }
}

```


### Bacteria

(Everything below is the same process as above)

Same as in NMR_Findings_Initial, but ignoring more uninteresting chemicals. 
```{r}
# OTU table for fungi
# Only worry about genera
dfbacteria <- tax_glom(dfbacteria, taxrank = "Genus")

otu.bacte <- otu_table(dfbacteria)
otu.bacte <- otu.bacte[, intersect(rownames(nmr.meta), colnames(otu.bacte))]
dim(otu.bacte)

otu.bacte <- sweep(otu.bacte, 2, colSums(otu.bacte), "/")
# Require that at each taxon is present at at least 5% in at least 3 samples
otu.bacte <- t(na.exclude(otu.bacte[rowSums(otu.bacte > 0.001) > 2, ]))
nmr.meta.bacteria <- nmr.meta[intersect(rownames(nmr.meta), rownames(otu.bacte)), 15:42]
nmr.meta.bacteria <- subset(nmr.meta.bacteria, select = c(
  -X2.Hydroxyisobutyrate,
  -X2.Hydroxyisovalerate,
  -X4.Guanidinobutanoate,
  -Caffeine,
  -Glucuronic.acid,
  -Glycerol,
  -Citraconate
    )
  )
mean_normalize <- function(vec) {
  return(vec / mean(vec))
}
nmr.meta.bacteria[, 2:dim(nmr.meta.bacteria)[2]] <- apply(
  nmr.meta.bacteria[, 2:dim(nmr.meta.bacteria)[2]], 
  2, 
  mean_normalize
  )

```

For each taxon, make a model using all the chemicals as explanatory variables, then use stepwise regression to find a stronger model. This code automatically does FDR correction for the variables included, but DOES NOT apply necessary corrections for the model selection process. 
```{r}
tax_df <- data.frame(tax_table(dfbacteria))

pvals <- c()
for (species in colnames(otu.bacte)) {
  model.initial <- lm(otu.bacte[, species] ~ Age..days. : ., data = nmr.meta.bacteria)
  #model.final <- summary(model.initial)
  model.final <- summary(step(model.initial, trace = FALSE))
  pvals <- append(pvals, model.final$coefficients[, 4])
  #adjust_p_values_multiple(models)
}
pvals_adj <- p.adjust(pvals)
for (species in colnames(otu.bacte)) {
  model.initial <- lm(otu.bacte[, species] ~ Age..days. : ., data = nmr.meta.bacteria)
  #model.final <- summary(model.initial)
  model.final <- summary(step(model.initial, trace = FALSE))
  n <- length(model.final$coefficients[, 4])
  model.final$coefficients[, 4] <- pvals_adj[1:n]
  pvals_adj < pvals_adj[n + 1:length(pvals)]
  print(paste(species, " (", tax_df[species, "Genus"], "):", sep = ""))
  print(model.final)
  #print(pvals_adj)
  #print(n)
}

```

Create a linear model using one chemical as the response variable and all others as explanatory variables. Then plot the residuals vs. some species of interest. 
```{r}
plot_resid_bacteria <- function(taxon, chemical){
  tax_df <- data.frame(tax_table(dfbacteria))
  genus <- tax_df[taxon, "Genus"]
  lm(nmr.meta.bacteria[, chemical] ~ ., data = nmr.meta.bacteria[colnames(nmr.meta.bacteria) != chemical]) -> lm1
  step(lm1, trace = FALSE) -> lm1
  plot_title <- paste(genus, "vs", chemical, "residuals")

  ggplot(mapping = aes(
    otu.bacte[, taxon],
    lm1$residuals)) +
    geom_point() +
    geom_smooth(method = "lm") +
    xlab(taxon) +
    ylab(chemical) +
    ggtitle(plot_title)
    
    if (askYesNo("Save plot?")){
      ggsave(paste("~/Documents/Classes/MB 599/Plots/", plot_title, ".png", sep = ""))
  }
}

```

Create a simple plot of a fungal species vs a chemical. 
```{r}
plot_bacteria <- function(taxon, chemical){
  tax_df <- data.frame(tax_table(dfbacteria))
  genus <- tax_df[taxon, "Genus"]
  lm(nmr.meta.bacteria[, chemical] ~ ., data = nmr.meta.bacteria[colnames(nmr.meta.bacteria) != chemical]) -> lm1
  step(lm1, trace = FALSE) -> lm1
  plot_title <- paste(genus, "vs", chemical)

  ggplot(mapping = aes(
    otu.bacte[, taxon],
    nmr.meta.bacteria[, chemical])) +
    geom_point() +
    geom_smooth(method = "lm") +
    xlab(taxon) +
    ylab(chemical) +
    ggtitle(plot_title)

  if (askYesNo("Save plot?")){
    ggsave(paste("~/Documents/Classes/MB 599/Plots/", plot_title, ".png", sep = ""))
  }
}
# test the function.
ggplot(mapping = aes(
    otu.bacte[, "ASV127"],
    nmr.meta.bacteria[, "Gluconic.acid"])) +
    geom_point() +
    geom_smooth(method = "lm")

plot_bacteria("ASV127", "Acetic.Acid")

```