---
title: "bacteria phyloseq"
output: html_document
---

```{r setup, include=FALSE}
Load R packages 
```{r setup, include=FALSE}
library(plotly)
library(phyloseq)
library(devtools)
library(dplyr)
library(lme4)
library(emmeans)
library(lme4)
library(emmeans)
library(ggplot2)
library(leaps)
library(car)
library(ape)
```


Load phyloseq objects, previously made by KH
```{r}
dfbacteria <- readRDS("~/Documents/Data/HAMS/phyloseq /dfbacteria.rds") #load the bacteria phyloseq object
dfbacteria = filter_taxa(dfbacteria, function(x) sum(x > 3) > (0.20*length(x)), TRUE) ##Remove taxa not seen more than 3 times in at least 20% of the samples. This protects against an OTU with small mean & trivially large C.V.
dfbacteria <- tax_glom(dfbacteria, taxrank = "Genus") #merge genus-level
```


Transform ASV table to relative abundance to look at composition of bacteria and fungi at a 
```{r}
bact_relative_abun<- transform_sample_counts(dfbacteria, function(x) x / sum(x) )
```

```{r}
random_tree = rtree(ntaxa(dfbacteria), rooted=TRUE, tip.label=taxa_names(dfbacteria))
plot(random_tree)
dfbacteria = merge_phyloseq(dfbacteria, random_tree)

# OTU table for bacteria
# Only worry about genera for right now
tax_table(bact_relative_abun)[, "Genus"] #Need to remove unassigned
# exclude variables unassigned 
dfbact.genus <- subset_taxa(bact_relative_abun, Genus != "Unassigned")

otu.bac <- otu_table(dfbact.genus) #create a data frame of RA
head(otu.bac)

# Convert to sample proportions. 
#otu.fungi = as.data.frame(t(otu.fungi))
#total = rowSums(otu.fungi)
#total #total read count
#otu.fungi <- otu.fungi / total 
#head(otu.fungi)
#data = otu.fungi@.Data
```



```{r}
#Phyloseq object with PMAxx 
dfbact.PMAxx <- subset_samples(dfbact.genus, Source == "Fermentation Batch")

#Phyloseq object - SCOBY-based without PMAxx
dfbact.genus <- subset_samples(dfbact.genus, Sample.Description == "Primary"& PMAxx. == "No")

dfbacteria@sam_data$Timepoint..Day. = as.numeric(dfbacteria@sam_data$Timepoint..Day.)
dfbacteria <- subset_samples(dfbacteria,Timepoint..Day. <= 7)

```


alpha diversity
```{r}
dfbact.genus@sam_data$Timepoint..Day. = as.numeric(dfbact.genus@sam_data$Timepoint..Day.)
dfbact.genus <- subset_samples(dfbact.genus,Timepoint..Day. <= 7)
sample_data(dfbact.genus)$shannon.physeq <- estimate_richness(dfbact.genus, measures="Shannon")
sample_data(dfbact.genus)$simpson.physeq <- estimate_richness(dfbact.genus, measures="Simpson")
plot_richness(dfbact.genus, "Timepoint..Day.", measures="Simpson", color="Sample.Type") + theme_bw()

```



```{r}
Alpha diversity comparison- PMAxx and non-PMAxx
```{r}
sample_data(dfbact.PMAxx)$simpson.physeq <- estimate_richness(dfbact.PMAxx, measures="Simpson")
dfbact.PMAxx@sam_data$Timepoint..Day. = as.numeric(dfbact.PMAxx@sam_data$Timepoint..Day.)
plot_richness(dfbact.PMAxx, "Timepoint..Day.", measures="Simpson", color="PMAxx.") + theme_bw()
```



Format simpson dataframe
```{r}
simpson = dfbact.genus@sam_data$simpson.physeq
write.csv(simpson,"~/Documents/Data/HAMS/phyloseq /bacsimpson.csv", row.names = TRUE) #export to add replicate ID names

#load dataset
#simpson_bacteria <- read_csv("Documents/Data/HAMS/phyloseq /simpson.bacteria.csv")"
simp.bact = simpson_bacteria
head(simp.bact)
class(simp.bact$Timepoint)
lmeModel = lmer(Simpson ~ SampleType*Timepoint + (1|ID), data=simp.bact)
anova(lmeModel, type='marginal')
```

Post-hoc for linear mixed model
```{r}
emmeans(lmeModel, pairwise ~ SampleType)
```

Format Simpson Comparison for PMAxx and non-PMAxx
```{r}
sample_data(dfbact.PMAxx)$simpson.physeq <- estimate_richness(dfbact.PMAxx, measures="Simpson")
simpson.pmaxx.b = dfbact.PMAxx@sam_data$simpson.physeq
write.csv(simpson.pmaxx.b,"~/Documents/Data/HAMS/phyloseq /bacteriasimpsonPMAxx.csv", row.names = TRUE) #export to add replicate ID names

#load dataset
#bacteriasimpsonPMAxx <- read_csv("bacteriasimpsonPMAxx.csv")
simp.pmaxx = bacteriasimpsonPMAxx
head(simp.pmaxx)
class(simp.pmaxx$Timepoint)
lmeModel = lmer(Simpson ~ Timepoint + (1|ID), data=simp.pmaxx)
anova(lmeModel, type = "marginal")
summary(lmeModel)

simps_bact <- lm(Simpson~PMAxx*Timepoint+SampleType, data=simp.pmaxx)
anova(simps_bact)

```


```{r}
dfbact.genus@sam_data$Timepoint..Day. = as.character(dfbact.genus@sam_data$Timepoint..Day.)
plot_bar(dfbact.genus, x="Timepoint..Day.", fill = "Genus") + geom_bar(aes(color=Genus, fill=Genus), stat="identity", position="stack") +  facet_wrap(~Sample.Type)
```

Format Export otutable to make linear models 
```{r}
otu.bacteria <- otu_table(dfbact.genus)
write.csv(otu.bacteria@.Data,"~/Documents/Data/HAMS/phyloseq /bactOTU.csv", row.names = TRUE)
otu_bacteria <- read_csv("Documents/Data/HAMS/phyloseq /otu.bacteria.csv", 
+     col_types = cols(X7 = col_skip(), X8 = col_skip(), 
+         X9 = col_skip(), X10 = col_skip(), 
+         X11 = col_skip(), X12 = col_skip(), 
+         X13 = col_skip()))
```

Build regression plots
```{r}
```{r}
bacteriaOTU = otu_bacteria
head(otu_bacteria)
tax_table(dfbact.genus)[, "Genus"] #Need to remove unassigned

# ASV076 "Lactobacillus"    ASV125 "Acetobacter"      ASV127 "Gluconacetobacter"
#ASV129 "Tanticharoenia"  

# ASV076 "Lactobacillus"
p = ggplot(bacteriaOTU, aes(x= Timepoint, y= ASV076, shape = SampleType, color = SampleType)) +     geom_point() + 
  scale_color_manual(values=c('lightsteelblue', 'mediumblue')) +
  theme_classic() +
  geom_smooth(method=lm) 
p

#ASV125 "Acetobacter" 
p = ggplot(bacteriaOTU, aes(x= Timepoint, y= ASV125, shape = SampleType, color = SampleType)) +     geom_point() + 
  scale_color_manual(values=c('lightsteelblue', 'mediumblue')) +
  theme_classic() +
  geom_smooth(method=lm) 
p

# ASV127 "KOMAGATEIBACTER"
p = ggplot(bacteriaOTU, aes(x= Timepoint, y= ASV127, shape = SampleType, color = SampleType)) +     geom_point() + 
  scale_color_manual(values=c('lightsteelblue', 'mediumblue')) +
  theme_classic() +
  geom_smooth(method=lm) 
p

#ASV129 "Tantechorenia"
p = ggplot(bacteriaOTU, aes(x= Timepoint, y= ASV129, shape = SampleType, color = SampleType)) +     geom_point() + 
  scale_color_manual(values=c('purple', 'darkpurple')) +
  theme_classic() +
  geom_smooth(method=lm) 
p

```



Build ordinate plots
```{r}
## to do stilllllll
```

Use AIC to determine best linear model for major genera
```{r}

otu_best_ASV076<- regsubsets(ASV076~., bacteriaOTU, nbest=10, really.big=T)
ASV076_best <- summary(otu_best_ASV076)
p <- rowSums(ASV076_best$which)
models <- with(ASV076_best,data.frame(p, rss, adjr2, cp, bic))
which.min(models$cp)
n <- nrow(fungusOTU)
models$aic <- with(models, n*log(rss/n)+2*(p+1))
best_aic <- which.min(models$aic)
ASV076_best$which[which.min(models$aic),]
```

```
```{r}
ASV76_model <- lm(ASV076~ASV125+ASV127+ASV129+Timepoint+SampleType, data=bacteriaOTU)
vif(ASV76_model)
anova(ASV76_model)
```


Use AIC to determine best linear model for major genera
```{r}

otu_best_ASV125<- regsubsets(ASV125~., bacteriaOTU, nbest=10, really.big=T)
ASV125_best <- summary(otu_best_ASV125)
p <- rowSums(ASV125_best$which)
models <- with(ASV125_best,data.frame(p, rss, adjr2, cp, bic))
which.min(models$cp)
n <- nrow(fungusOTU)
models$aic <- with(models, n*log(rss/n)+2*(p+1))
best_aic <- which.min(models$aic)
ASV125_best$which[which.min(models$aic),]
```

```
```{r}
ASV125_model <- lm(ASV125~ASV076+ASV127+ASV129+Timepoint, data=bacteriaOTU)
vif(ASV125_model)
anova(ASV125_model)
```

Use AIC to determine best linear model for major genera
```{r}

otu_best_ASV127<- regsubsets(ASV127~., bacteriaOTU, nbest=10, really.big=T)
ASV127_best <- summary(otu_best_ASV127)
p <- rowSums(ASV127_best$which)
models <- with(ASV127_best,data.frame(p, rss, adjr2, cp, bic))
which.min(models$cp)
n <- nrow(fungusOTU)
models$aic <- with(models, n*log(rss/n)+2*(p+1))
best_aic <- which.min(models$aic)
ASV127_best$which[which.min(models$aic),]
```

```
```{r}
ASV127_model <- lm(ASV127~ASV076+ASV125+ASV129+Timepoint+SampleType, data=bacteriaOTU)
vif(ASV127_model)
anova(ASV127_model)
```

Use AIC to determine best linear model for major genera
```{r}

otu_best_ASV129<- regsubsets(ASV129~., bacteriaOTU, nbest=10, really.big=T)
ASV129_best <- summary(otu_best_ASV129)
p <- rowSums(ASV129_best$which)
models <- with(ASV129_best,data.frame(p, rss, adjr2, cp, bic))
which.min(models$cp)
n <- nrow(fungusOTU)
models$aic <- with(models, n*log(rss/n)+2*(p+1))
best_aic <- which.min(models$aic)
ASV129_best$which[which.min(models$aic),]
```

```
```{r}
ASV129_model <- lm(ASV129~ASV076+ASV125+ASV127+Timepoint+SampleType, data=bacteriaOTU)
vif(ASV129_model)
anova(ASV129_model)
```
Buikld ordinate plots and compare diversity using unifrac distance
```{r}
wUF.ordu = ordinate(dfbacteria, method="NMDS", distance="unifrac", weighted=TRUE)
dfbacteria.ord <- ordinate(dfbacteria, "NMDS", "bray")
par(mfrow=c(1,1))

plot(dfbacteria.ord)

wunifrac_dist = phyloseq::distance(dfbacteria, method="unifrac", weighted=T)
ordination = ordinate(dfbacteria, method="PCoA", distance=wunifrac_dist)
plot_ordination(dfbacteria, ordination, color="PMAxx.") + theme(aspect.ratio=1) + 
  scale_color_manual(values=c('darkgoldenrod2', 'cornsilk4')) + theme_bw()
  
adonis(wunifrac_dist ~ sample_data(dfbacteria)$PMAxx.)  

```

