---
title: "Phyloseq Work flow for ITS data for Microbiota Succession Analysis"
output: html_document
---

Load R packages 
```{r setup, include=FALSE}
library(plotly)
library(phyloseq)
library(devtools)
library(ggbiplot)
library(dplyr)
library(lme4)
library(emmeans)
library(ggplot2)
library(leaps)
library(car)
library(vegan)
install.packages("remotes")
remotes::install_github("vmikk/metagMisc")
library(remotes)
library(meta)
```


Load phyloseq objects, previously made by KH
```{r}
dffungi <- readRDS("~/Documents/Data/HAMS/phyloseq /dffungi.rds") #load the fungal phyloseq object 
dffungi = filter_taxa(dffungi, function(x) sum(x > 3) > (0.20*length(x)), TRUE) ##Remove taxa not seen more than 3 times in at least 20% of the samples. This protects against an OTU with small mean & trivially large C.V.
dffungi <- tax_glom(dffungi, taxrank = "Genus") #merge genus-level
random_tree = rtree(ntaxa(dffungi), rooted=TRUE, tip.label=taxa_names(dffungi))
plot(random_tree)
dffungi = merge_phyloseq(dffungi, random_tree)
dffungi = phyloseq_rm_na_tax(dffungi)
```


Transform ASV table to relative abundance to look at composition of bacteria and fungi at a 
```{r}
fung_relative_abun<- transform_sample_counts(dffungi, function(x) x / sum(x) )
```

```{r}
# OTU table for fungi
# Only worry about genera for right now
tax_table(fung_relative_abun)[, "Genus"] #Need to remove unassigned
# exclude variables unassigned 
dffungi.genus <- subset_taxa(fung_relative_abun, Genus != "Unassigned")

otu.fungi <- otu_table(dffungi.genus) #create a data frame of RA
head(otu.fungi)

# Convert to sample proportions. 
#otu.fungi = as.data.frame(t(otu.fungi))
#total = rowSums(otu.fungi)
#total #total read count
#otu.fungi <- otu.fungi / total 
#head(otu.fungi)
#data = otu.fungi@.Data
```



```{r}
#Phyloseq object with PMAxx 
dffungi.PMAxx <- subset_samples(dffungi.genus, Source == "Fermentation Batch")
random_tree = rtree(ntaxa(dffungi.PMAxx), rooted=TRUE, tip.label=taxa_names(dffungi.PMAxx))
plot(random_tree)
dffungi.PMAxx = merge_phyloseq(dffungi.PMAxx, random_tree)

#Phyloseq object - SCOBY-based without PMAxx
dffungi.genus <- subset_samples(dffungi.genus, Sample.Description == "Primary"& PMAxx. == "No")
#Add random phylogenic tree
random_tree = rtree(ntaxa(dffungi.genus), rooted=TRUE, tip.label=taxa_names(dffungi.genus))
plot(random_tree)
physeq.tree = merge_phyloseq(physeq, NJ.tree)

```


alpha diversity
```{r}
dffungi.genus@sam_data$Timepoint..Day. = as.numeric(dffungi.genus@sam_data$Timepoint..Day.)
dffungi.genus <- subset_samples(dffungi.genus,Timepoint..Day. <= 7)
sample_data(dffungi.genus)$shannon.physeq <- estimate_richness(dffungi.genus, measures="Shannon")
sample_data(dffungi.genus)$simpson.physeq <- estimate_richness(dffungi.genus, measures="Simpson")
plot_richness(dffungi.genus, "Timepoint..Day.", measures="Simpson", color="Sample.Type") + theme_bw()

```

Alpha diversity comparison- PMAxx and non-PMAxx
```{r}
head(dffungi@sam_data)
sample_data(dffungi.PMAxx)$simpson.physeq <- estimate_richness(dffungi.PMAxx, measures="Simpson")
dffungi.PMAxx@sam_data$Timepoint..Day. = as.numeric(dffungi.PMAxx@sam_data$Timepoint..Day.)
plot_richness(dffungi.PMAxx, "Timepoint..Day.", measures="Simpson", color="PMAxx.") + theme_bw()
```



Format simpson dataframe for timepoint analysis
```{r}
sample_data(dffungi.genus)$simpson.physeq <- estimate_richness(dffungi.genus, measures="Simpson")
simpson = dffungi.genus@sam_data$simpson.physeq
write.csv(simpson,"~/Documents/Data/HAMS/phyloseq /fungsimpson.csv", row.names = TRUE) #export to add replicate ID names
simpson = dffungi.genus@sam_data$simpson.physeq


#load dataset
#simpson_fungi <- read_csv("Documents/Data/HAMS/phyloseq /simpson.fungi.csv", 
#+     col_types = cols(X12 = col_skip()))
simp.fungi = simpson_fungi
head(simp.fungi)
class(simp.fungi$Timepoint)
lmeModel = lmer(Simpson ~ SampleType*Timepoint + (1|ID), data=simp.fungi)
anova(lmeModel, type = "marginal")


```

Format simpson dataframe for timepoint analysis
```{r}
sample_data(dffungi.PMAxx)$simpson.physeq <- estimate_richness(dffungi.PMAxx, measures="Simpson")
simpson.pmaxx = dffungi.PMAxx@sam_data$simpson.physeq
pma.samps = dffungi.PMAxx@sam_data
write.csv(simpson.pmaxx,"~/Documents/Data/HAMS/phyloseq /fungisimpsonPMAxx.csv", row.names = TRUE) #export to add replicate ID names
write.csv(pma.samps,"~/Documents/Data/HAMS/phyloseq /fungisampsPMAxx.csv", row.names = TRUE) #export to add replicate ID names

#load dataset
#fungi_PMAxx <- read_csv("fungisimpsonPMAxx.csv")
simp.pmaxx = fungisimpsonPMAxx
head(simp.pmaxx)
class(simp.pmaxx$Timepoint)
lmeModel = lmer(Simpson ~ SampleType*Timepoint + (1|ID), data=simp.pmaxx)
anova(lmeModel, type = "marginal")

simps_fung <- lm(Simpson~PMAxx*Timepoint+SampleType, data=simp.pmaxx)
anova(simps_fung)

```


```{r}
plot_bar(dffungi.genus, x="Timepoint..Day.", fill = "Genus") + geom_bar(aes(color=Genus, fill=Genus), stat="identity", position="stack") +  facet_wrap(~Sample.Type)
```

Format Export otutable to make linear models 
```{r}
otu.fungi <- otu_table(dffungi.genus)
head(otu.fungi)
write.csv(otu.fungi@.Data,"~/Documents/Data/HAMS/phyloseq /fungusOTU.csv", row.names = TRUE)
```


Build regression plots
```{r}
```{r}
head(fungusOTU)
tax_table(dffungi.genus)[, "Genus"] #Need to remove unassigned
#ASV030 "Brettanomyces", ASV034 "Issatchenkia", ASV036 "Kazachstania", ASV038 "Lachancea", ASV048 "Starmerella" 

#ASV030 Brettanomyces
p = ggplot(fungusOTU, aes(x= Timepoint, y= ASV030, shape = SampleType, color = SampleType)) +     geom_point() + 
  scale_color_manual(values=c('green', 'darkgreen')) +
  theme_classic() +
  geom_smooth(method=lm) 
p

#ASV034 "Issatchenkia"
p = ggplot(fungusOTU, aes(x= Timepoint, y= ASV034, shape = SampleType, color = SampleType)) +     geom_point() + 
  scale_color_manual(values=c('green', 'darkgreen')) +
  theme_classic() +
  geom_smooth(method=lm) 
p

#ASV036 "Kazachstania"
p = ggplot(fungusOTU, aes(x= Timepoint, y= ASV036, shape = SampleType, color = SampleType)) +     geom_point() + 
  scale_color_manual(values=c('green', 'darkgreen')) +
  theme_classic() +
  geom_smooth(method=lm) 
p

#ASV038 "Lachancea"
p = ggplot(fungusOTU, aes(x= Timepoint, y= ASV038, shape = SampleType, color = SampleType)) +     geom_point() + 
  scale_color_manual(values=c('green', 'darkgreen')) +
  theme_classic() +
  geom_smooth(method=lm) 
p

#ASV048 "Starmerella"
p = ggplot(fungusOTU, aes(x= Timepoint, y= ASV048, shape = SampleType, color = SampleType)) +     geom_point() + 
  scale_color_manual(values=c('green', 'darkgreen')) +
  theme_classic() +
  geom_smooth(method=lm) 
p
```

Build best of fit linear models
```{r}
fungusOTU <- read_csv("fungusOTU.csv")
head(fungusOTU)
otu_best_ASV030 <- regsubsets(ASV030~., fungusOTU, nbest=10, really.big=T)
ASV030_best <- summary(otu_best_ASV030)
p <- rowSums(ASV030_best$which)
models <- with(ASV030_best,data.frame(p, rss, adjr2, cp, bic))
which.min(models$cp)
n <- nrow(fungusOTU)
models$aic <- with(models, n*log(rss/n)+2*(p+1))
best_aic <- which.min(models$aic)
ASV030_best$which[which.min(models$aic),]
```
```{r}
ASV30_model <- lm(ASV030~ASV034+ASV038+ASV048+SampleType, data=fungusOTU)
vif(ASV30_model)
anova(ASV30_model)
```





```{r}
otu_best_ASV034<- regsubsets(ASV034~., fungusOTU, nbest=10, really.big=T)
ASV034_best <- summary(otu_best_ASV034)
p <- rowSums(ASV034_best$which)
models <- with(ASV034_best,data.frame(p, rss, adjr2, cp, bic))
which.min(models$cp)
n <- nrow(fungusOTU)
models$aic <- with(models, n*log(rss/n)+2*(p+1))
best_aic <- which.min(models$aic)
ASV034_best$which[which.min(models$aic),]
```

```{r}
ASV34_model <- lm(ASV034~ASV030+ASV038+ASV048+SampleType, data=fungusOTU)
vif(ASV30_model)
anova(ASV34_model)
```


```{r}
otu_best_ASV036<- regsubsets(ASV036~., fungusOTU, nbest=10, really.big=T)
ASV036_best <- summary(otu_best_ASV036)
p <- rowSums(ASV036_best$which)
models <- with(ASV036_best,data.frame(p, rss, adjr2, cp, bic))
which.min(models$cp)
n <- nrow(fungusOTU)
models$aic <- with(models, n*log(rss/n)+2*(p+1))
best_aic <- which.min(models$aic)
ASV036_best$which[which.min(models$aic),]
```
```{r}
```{r}
ASV36_model <- lm(ASV036~ASV048, data=fungusOTU)
vif(ASV36_model)
anova(ASV36_model)
```

```

```{r}
otu_best_ASV038 <- regsubsets(ASV038~., fungusOTU, nbest=10, really.big=T)
ASV038_best <- summary(otu_best_ASV038)
p <- rowSums(ASV038_best$which)
models <- with(ASV038_best,data.frame(p, rss, adjr2, cp, bic))
which.min(models$cp)
n <- nrow(fungusOTU)
models$aic <- with(models, n*log(rss/n)+2*(p+1))
best_aic <- which.min(models$aic)
ASV038_best$which[which.min(models$aic),]
```

```{r}
ASV38_model <- lm(ASV038~ASV048+ASV030+ASV034, data=fungusOTU)
vif(ASV38_model)
anova(ASV38_model)
```



```{r}
```
```{r}
otu_best_ASV048 <- regsubsets(ASV048~., fungusOTU, nbest=10, really.big=T)
ASV048_best <- summary(otu_best_ASV048)
p <- rowSums(ASV048_best$which)
models <- with(ASV048_best,data.frame(p, rss, adjr2, cp, bic))
which.min(models$cp)
n <- nrow(fungusOTU)
models$aic <- with(models, n*log(rss/n)+2*(p+1))
best_aic <- which.min(models$aic)
ASV048_best$which[which.min(models$aic),]
```

```{r}
ASV48_model <- lm(ASV048~ASV038+ASV030+ASV034, data=fungusOTU)
vif(ASV48_model)
anova(ASV48_model)
```

Create ordinate plot to compare PMAxx treated and non-treated samples 
```{r}
wUF.ordu = ordinate(dffungi, method="NMDS", distance="unifrac", weighted=TRUE)
fungi.ord <- ordinate(dffungi, "NMDS", "bray")
par(mfrow=c(1,1))

plot(fungi.ord)

wunifrac_dist = phyloseq::distance(dffungi, method="unifrac", weighted=F)
ordination = ordinate(dffungi, method="PCoA", distance=wunifrac_dist)
plot_ordination(dffungi, ordination, color="PMAxx.") + theme(aspect.ratio=1) + 
  scale_color_manual(values=c('darkgoldenrod2', 'cornsilk4')) + theme_bw()
  
adonis(wunifrac_dist ~ sample_data(dffungi)$PMAxx.)  
```

